---
title: "Chemicals In Strawberry Farms"
author: "Ruxin Liu"
date: "10/20/2020"
output: ioslides_presentation
---

```{r , include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)
unfood_table <- read.csv("Unfood.csv")
```

## Berry Data

- Collected by the United States Department of Agriculture (USDA)
  Available from: https://quickstats.nass.usda.gov/ 
  
- 13238 rows and 21 columns

- Contained information about 3 kinds of berries: blueberries, raspberries & strawberries

- Year ranged from 2015 - 2019

- Goal: Compared the usage of synthetic chemicals applied for planting strawberries




## Data Cleaning & Variable Selection

- Removed 12 variables that only had one unique value, e.g "Program", "Week Ending",
"Geo Level" and etc

- Separated "Data Item" into "berry" (types of berry), "market" (marketing condition), 
"harvest" (harvest condition) and "units" (units of values collected)

- Reorganized and compared "Domain" and "Domain Category" into "materials" and "chemicals

- 8 variables left after data cleaning: Year, State, market, harvest, materials, 
chemical, Values and units

- Removed all NA values and D values 

- Created sub-dataset that showed the pounds of chemicals usage per acre per 
application for strawberries
 
  

# Exploratory Data Analysis (EDA)

## Difference among states
- Continuous and relatively large use of chemicals California and Florida

```{r, echo = FALSE, warning = FALSE, message = FALSE}
unfood_table %>% 
ggplot(aes(x = Year)) +
  geom_bar(aes(fill = State)) +
  ggtitle("Number Of Strawberry Farms That Used Chemicals")
```

## Difference among chemical types
- Herbicides and insecticides were used the most 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
unfood_table %>% 
ggplot(aes(x = Year)) +
  geom_bar(aes(fill = chemical)) +
  ggtitle("The Annual Usage Frequency Of Each Chemicals")
```

## Fungicide

- Florida on average had a higher average pounds of fungicide usage per acre per
application.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
fungicide <- unfood_table %>% 
  filter(chemical == "FUNGICIDE") %>% 
  group_by(Year, State) %>% 
  summarise(min = min(as.numeric(Value)), 
            max = max(as.numeric(Value)),
            mean = mean(as.numeric(Value)), 
            sum = sum(as.numeric(Value)))
knitr::kable(fungicide, digits = 3) %>% 
  kableExtra::kable_styling(latex_options = "hold_position")
```

## Herbicide
- California used herbicides continuously with relatively high level.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
herbicide <- unfood_table %>% 
  filter(chemical == "HERBICIDE") %>% 
  group_by(Year, State) %>% 
  summarise(min = min(as.numeric(Value)), 
            max = max(as.numeric(Value)),
            mean = mean(as.numeric(Value)), 
            sum = sum(as.numeric(Value)))
knitr::kable(herbicide, caption = "Herbicide Application ", digits = 3) %>% 
  kableExtra::kable_styling(latex_options = "hold_position")
```

## Insecticide

- California always had the highest mean usage of insecticide compared to other
states. 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
insecticide <- unfood_table %>% 
  filter(chemical == "INSECTICIDE") %>% 
  group_by(Year, State) %>% 
  summarise(min = min(as.numeric(Value)), 
            max = max(as.numeric(Value)),
            mean = mean(as.numeric(Value)), 
            sum = sum(as.numeric(Value)))
knitr::kable(insecticide, caption = "Insecticide Application ", digits = 3) %>% 
  kableExtra::kable_styling(latex_options = "hold_position")
```


## Conclusion 

- The pounds of usage per acre per chemical application differed by the type of
chemicals and also the years.

- For fungicides, herbicides and insecticides, the maximum usage per pound per 
acre per application all happened in California. 

- Suggests that there might be issues with the strawberry industry in California.

- Large amount of harmful pesticide usage -- already an existing and noticed problem 
in California

- Maybe consider about organic strawberries?


## Reference & Thank you for listening!
<font size="3">
- R Packages

<font size="3"> 1. Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software,
4(43),1686, https://doi.org/10.21105/joss.01686

<font size="3"> 2.  Stefan Milton Bache and Hadley Wickham (2014). magrittr: A Forward-Pipe Operator
for R. R package version 1.5. https://CRAN.R-project.org/package=magrittr

<font size="3"> 3. H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New 
York, 2016.

<font size="3"> 4. Hao Zhu (2019). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R
  package version 1.1.0. https://CRAN.R-project.org/package=kableExtra
  
- Bibliography

<font size="3"> 1. MA615 notes (class 14 - class 17)

<font size="3"> 2. USDA (2019). *Quick Stats* [online]. Available from:
https://quickstats.nass.usda.gov/results/D416E96E-3D5C-324C-9334-1D38DF88FFF1
[accessed 16 October 2020].

<font size="3"> 3. Smithsonian MAGAZINE (2018). *The Toxic Rise of the California Strawberry* 
[online]. Available from:
https://www.smithsonianmag.com/science-nature/toxic-rise-california-strawberry-180967934/
[accessed 18 October 2020].

<font size="3"> 4. The Guardian (2014). *California’s strawberry industry is hooked on dangerous pesticides* 
[online]. Available from:
https://www.theguardian.com/us-news/2014/nov/10/-sp-california-strawberry-industry-pesticides [accessed 18 October 2020].





